manuscript_outline

### 2025-06-11 docs/manuscript/paper.Rmd
add ecological section with karraker data

library(hatchR)
library(tidyverse)
library(patchwork)
```

@@

Recreater @karraker2006 with data from @siegel2023

@@

@Tang example using females

>>>>

library(hatchR)
library(tidyverse)
library(patchwork)
library(ggridges)
library(modelbased)
```

@@

Recreate @karraker2006 with data from @siegel2023

```{r, eval=FALSE}
ascaphus_sites <- readxl::read_xlsx("../../data/ascaphus_locations.xlsx") |> drop_na(comid)

#########################################
# ONLY RUN IF YOU DOWNLOAD SIEGEL FILES #
#########################################

# print huc_6 to download from https://zenodo.org/records/8174951

huc_6 <-ascaphus_sites |> 
  drop_na(comid) |> 
  pull(HUC_6) |> 
  unique()

ascaphus_siegel_data <- NULL

for( h in huc_6 ){
huc_6_files <- fs::dir_ls(paste0("../../data/siegel_data/st_pred_", h))

#filter to huc10 to subset files
huc_10 <- ascaphus_sites |> 
  filter(HUC_6 == h) |> 
  pull(HUC_10) |> 
  unique()

# get comids to filter mega df created below
comids <- ascaphus_sites |> 
  drop_na(comid) |> 
  filter(HUC_10 %in% huc_10) |> 
  pull(comid) |> 
  unique()

# change to single character for pattern in function below
huc_10_patterns <- ifelse(length(huc_10) >1, paste0(huc_10, collapse = "|"), paste0(huc_10))

# put files in a df together
huc_10_df <- huc_6_files |>
  str_subset(pattern = (huc_10_patterns)) |> # str subset on huc10
  map_df(read_csv, .id = "file", col_types = cols()) |> 
  filter(COMID %in% comids)

#append to single file
ascaphus_siegel_data <- rbind(ascaphus_siegel_data, huc_10_df)

}

write_rds( ascaphus_siegel_data, "../../data/ascaphus_siegel_data.rds", compress = "gz" )
```

\

```{r, fig.height= 6 , fig.width= 8, fig.align = 'center', fig.cap="Estimated marginal means of days to hatch for coastal tailed frogs from various mountain ranges across their range."}
ascaphus_siegel_data <- readRDS("../../data/ascaphus_siegel_data.rds") |> 
  mutate(date = ymd(tim.date))

ascaphus_sites <- readxl::read_xlsx("../../data/ascaphus_locations.xlsx") |> drop_na(comid)

#karraker_data <- readxl::read_xls("../../data/ASMO_ASTR_egg_data.xls")

spawn_dates <- map(
  c(1991:2020), # year vector to map for custom function
  function(year) { # custom function to make dates
      paste0(year, "-07-24")
    }
  ) |> 
  unlist()

pred_hatch = NULL
for(sites in unique(ascaphus_siegel_data$COMID)) {

  site <-  ascaphus_siegel_data |> filter( COMID == sites) |> drop_na(prd.stream_temp)

 # spawn_dates <- spawn_dates[which(spawn_dates %in% site$tim.date)]

  hatch <- map(spawn_dates, 
               predict_phenology,
              data = site,
              temperature = prd.stream_temp,
              dates = date,
              model = ascaphus_mod$expression$expression) |> 
    map_dbl("days_to_develop")

  temp_df <- data.frame(days = hatch,
                        comid = rep(sites, times =length(hatch)),
                        year = year(spawn_dates))

  pred_hatch <- rbind(pred_hatch, temp_df)
}

pred_hatch <- pred_hatch |> 
  left_join(ascaphus_sites |> group_by(comid) |> slice_head( n =1), by = join_by(comid == comid))

# pred_hatch |> 
#   mutate(comid = as.factor(comid)) |> 
#   ggplot() +
#   geom_density_ridges(aes(y = year, x = days + ymd("2001-07-14"), fill = comid)) +
#   theme_classic() +
#   theme(legend.position = "none")
# 
# pred_hatch |> 
#   mutate(comid = as.factor(comid)) |> 
#   ggplot(aes(x = year, y = days , color= Mountain, group = comid)) +
#   geom_point() +
#   #geom_line() +
#   geom_smooth(method = "lm", se = FALSE) +
#   labs(x= "Year", y = "Hatch Date") +
#   theme_classic() +
#   theme(legend.position = c(0.8, 0.75))

mod <- lm(days ~ year*Mountain, data = pred_hatch)
# sjPlot::tab_model(mod)

estimate_means(mod, c("year", "Mountain")) |> 
  plot() +
  #geom_point(data = pred_hatch, aes(x = year, y = days, color = Mountain), size = 0.5 , alpha = 0.25) +
  labs(x = "Year", y = "Mean of Days to Hatch") +
  theme_classic()
```

@@

@tang2017 example using females
